Refinement for open automata

### 2021-06-14 2021-Refinement/report.tex
comments on intro

\usepackage{xunicode}
\usepackage{fontspec}
\usepackage[hmargin=3cm]{geometry}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{unicode-math}
\usepackage{tikz}
\usetikzlibrary{external}
\usepackage[colorlinks]{hyperref}
\AtBeginDocument{\renewcommand\setminus{\smallsetminus}}
\usepackage{polyglossia}
\setmainlanguage{english}

@@

\section{Introduction}
The open automata model is used to give the semantic of open pNets \cite{2007.10770}.
Open pNets are not petri nets but an intermediate representation used in the VerCors project to perform verification \cite{henrio:01252323}.
Verification is the process of checking that an implementation satisfies a specification.
Open pNets with the open automaton semantics and equivalence relation introduced in previous work \cite{2007.10770} were sucessfully used to model and verify BIP and GCM \cite{qin:01823507, ameurboulifa:01526055}.
However equivalences relations are not always sufficient to perform verification, refinement relations are also used.
For instance an example of open pNet where equivalence is not sufficient was encountered in an article about algorithms fo weak bisimulation on open automata \cite{wang:03126313}.

The contributions of this article are the definition of composition for open automata without having to define open pNets, and the introduciton of several refinement relations for open automata. % TODO: expand

@@

We begin with Section \ref{sec:notations} by giving some notations used throughout this paper.
Then there are several sections dedicated to define all the interesting objects and properties:
In Section \ref{sec:def} we give a clean and up to date definition of open automata.
In Section \ref{sec:comp} we define their composition without requiring the use of open pNets.
In Section \ref{sec:proofelts} we adapt the standard properties related to refinement relation from LTS to open automata.
After that Section \ref{sec:prelref} is dedicated to progressively build a refinement relation.
Section \ref{sec:refinement} introduces and analyses the refinement relation on open automata.
% future work
% conclusion

@@

Families of values, or equivalently maps will be noted \(\mset{i \mapsto x_i}{i \in I}\), \(\mset{i \gets x_i}{i \in I}\) or \(x_i^{i \in I}\).
The latter can only be used if there is a generating expression; for instance \(\mpar{ax}^{x \in \setR}\) represents a scaling function, \(c^{i \in I}\) is a constant function over \(I\).
However \(\mbrc{\alpha \mapsto 1, \beta \mapsto 2, \gamma \mapsto 3}\) has no generating expression and is represented here with the finite version of first notation.

The disjoint union on set is noted \(\uplus\).
There are several ways of ensuring a union is disjoint, we will indifferently either suppose sets are disjoint or rename conflicting object (useful for variables).
The disjoint union of two maps \(\varphi: I \to X\) and \(\psi: J \to Y\) with \(I \cap J = \emptyset\) is \(\varphi \uplus \psi: I \uplus J \to X \cup Y\).

@@

The terms \(\terms\) is a term algebra.
As any term algebra it has constant symbols with arity, variables, and a typing mechanism to distinguish well-formed and ill-formed terms.
The action algebra \(\actions\) is another term algebra.
It can be a subset of the terms.
The formulas \(\formulas\) are at least the first order formulas over \(\terms\) and \(\actions\).
\end{defi}
The term algebras are arbitrary.
The formulas contain at least first order logic terms with an equality relation.
This equality relation is not necessarily a syntactic equality (\(2 + 2 \neq 4\) with a syntactic equality).

@@

Note that in the previous paragraph there are two relations:
The first is between states of two specific automata (\(R \subseteq S_1 \times S_2\)), the second is between automata (\(A_1 \leq A_2\)).
Generally the second is implicit because its definition is always \(\exists R \subseteq S_1 \times S_2, P\mpar{R}\) with \(P\) a property, and only talking about \(R\) and \(P\) is sufficient.
However later on in this article we will define more complex relations on automata so we clarify the distinction now.
A relation of second kind will be called ``relation on automata" whereas a relation of the first kind is called ``relation on configurations" as defined below.
In the work on FH-bisimulation \cite{henrio:01055091} such a relation \(R\) was defined as triples in \(S_1 \times S_2 \times \formulas\) where for \(\mpar{s_1, s_2} \in S_1 \times S_2\) there is a unique \(f \in \formulas\) such that \(\mpar{s_1, s_2, f} \in R\).
This is classically the definition of a function from \(S_1 \times S_2\) to \(\formulas\) so we will define it as such.
\begin{defi}[Relation on configurations]
A relation on configurations of \(\OAg[1]\) and \(\OAg[2]\) is a function \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\).

@@

The important properties we consider for the relation on states are being a simulation and the prevention of deadlocks as defined in the following definitions.
\begin{defi}[Pre-simulation]
Let \(A_1 \defobject \OAg[1]\) and \(A_2 \defobject \OAg[2]\).
A relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is a pre-simulation if it satisfies both
\defitem Initial configurations are related: \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\);
\defitem From related states, all out-transitions from \(A_1\) can be simulated in \(A_2\) and their target states are related:
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} t_1 \defobject \OTx{1}{}{1}{1} \in \fOT{s_1}, \forall \sigma: V_1 \uplus V_2 \uplus \fvars{t_1} \to \values, \\
	\mpar{\sigma \vdash R\mpar{s_1, s_2} \wedge g_1} \implies \bigsymb{\exists} t_2 \defobject \OTx{2}{}{2}{2} \in \fOT{s_2}, \exists \nu: \fvars{t_2} \to \values, \\
	\sigma \uplus \nu \vdash \alpha_1 = \alpha_2 \wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} \wedge g_2 \wedge R\mpar{s'_1, s'_2}\psubst{\psi_1 \uplus \psi_2}
\end{multline*}
\end{defi}
This definition translated to open automata the classical definition of simulation \cite{10.1007/3-540-54430-5_78}.
The last formula means that for every pair of related configurations of the automaton and every possible transitions (\(\vdash g_1\)) from the first automaton, there is a possible transition (\(\vdash g_2\)) such that the produced action matches (\(\alpha_1 = \alpha_2\)), the same holes' action (\(J'_1 \cap J'_2\)) matches (\(\beta_{1j} = \beta_{2j}\)) and the target states are related after variable update. % OR variableS
Apart from holes, this is a natural extension of the notion of simulation on LTS, which is the same definition without variables nor holes nor guards.
In FH-bisimulation, holes with the same name will receive the same open automaton when composed, and the resulting open automaton will still be equivalent.
Here we assume that one of the related open automata can be obtained by composing an open automaton in a hole, open automaton which can have holes itself.
This implies that there is no relation between their holes, for instance composing \(A_i\) in \(j_i\) (\(1 \leq i \leq n\)) of \(A_0\) gives the set of holes \(J \defobject \mpar{\biguplus_{0 \leq i \leq n} J_i} \setminus \mset{j_i}{1 \leq i \leq n}\) that is neither subset nor superset of \(J_0\).
Yet we still assume that in holes with the same name, the same open automaton will be composed.
Hence only the holes in common have to match their actions.
However it is not sufficient to imply properties on the relation on automatons like being a preorder.
It is only a necessary condition for any relation that will be defined with ``simulation" in its name.

@@

% Another way of understanding this predicate is that it is a fixpoint of the union of valuations (function \(f\) defined below) that contains the initial valuation:
% \begin{align*}
	% f\mpar{p} = \mbrc{\mbox{\nmm{s' \mapsto p\mpar{s'} \vee \bigsymb{\bigvee_\subbox{\OTg \,\in\, \fIT{s'}}} p\mpar{s}\psubst{\psi} \wedge g}}} && % TODO: fix that in some way
	% \sigma_0 \vdash \reach{A}\mpar{s_0} &&
	% f\mpar{\reach{A}} = \reach{A}
% \end{align*}
\begin{defi}[Non-locking composition, intuitive definition]
Let \(A_i \defobject \OAg[i], 0 \leq i \leq n\).
The composition \(A \defobject A_0\mdbrk{j_i \mapsto A_i \middle| 1 \leq i \leq n}\) is non-locking if \(A\) has a reachability predicate satisfying:
From reachable configurations, if there was a possible transition in \(A_0\) then there is a possible transition in \(A\).

Formally for \(A = \OAg\), it gives
\begin{multline*}
	\forall s \defobject \mpar{s_0, s'} \in S_0 \times \prod_\subbox{1 \leq i \leq n} S_i, \forall \sigma: V \to \values, \mpar{\sigma \vdash \reach{A}\mpar{s}} \implies \\
	\mpar{\exists t_0 \in \fOT{s_0}, \exists \nu_0: \fvars{t_0} \to \values, \sigma \uplus \nu_0 \vdash \fguard{t_0}} \implies \\
	\exists t \in \fOT{s}, \exists \nu: \fvars{t} \to \values, \sigma \uplus \nu \vdash \fguard{t}
\end{multline*}
\end{defi}
We can expand the alias \(V\) into \nmm{\biguplus_\subbox{1 \leq i \leq n} V_i \uplus V_0} and transform the last transition \(t\) into a pair \(\mpar{t, t_0}\) where the second transition is the one from \(A_0\) which generated \(t\).
Doing that makes the definition become mildly similar with the deadlock prevention one. % OR slightly
It is not a coincidence because their goal is to caracterise the same kind of compatibility between automata.
Actually it is possible to simplify both definition and make them even more similar.
\begin{defi}[Deadlock prevention, working definition]
A relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is deadlock reducing if it is a pre-simulation and it satisfies the following:
\[ \forall \mpar{s_1, s_2} \in S_1 \times S_2, V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \]
\end{defi}
The requirement here is that from any related configurations, if there is a possible transition in the second automaton (whatever the value of the free variables) then there is a possible transition in the first automaton (the existential quantifier is implicit in the notation \(\vdash\)).
The transition are not matched because the relation on configurations is a pre-simulation and it is sufficient to ensure it as will be shown in the proof.
Apart from the matching transitions this is exactly the same explaination as the intuitive definition.
The expected advantage of this formulation is that a SMT should behave better without the quantifiers.
Indeed the big disjunctions are not quantifiers because in practise the out-transitions are finitary, and so is the expansion of the disjunction. % Check if that construction is correct
\begin{lem}
The intuitive and working definition of deadlock prevention are equivalent.
\end{lem}
\begin{defi}[Non-locking composition, working definition]
\(A \defobject A_0\mdbrk{j_i \mapsto A_i \middle| 1 \leq i \leq n}\) is a non-locking composition if:
\[ \forall s \defobject \mpar{s_0, s'} \in S, V \uplus \biguplus_\subbox{t_0 \in \fOT{s_0}} \fvars{t_0} \vdash \reach{A}\mpar{s} \wedge \bigvee_\subbox{t_0 \in \fOT{s_0}} \fguard{t_0} \implies \bigvee_\subbox{t \in \fOT{s}} \fguard{t} \]
\end{defi}
\begin{lem}
The intuitive and working definition of non-locking composition are equivalent.
\end{lem}
An illustration of both the introduction of a deadlock by composition and the definition of non-locking composition is given in Appendix \ref{apx:lockcomp}.
Proof of both lemma are given in Appendix \ref{apx:lemeqd}.

This new definition of composition can replace the standard composition.
For the simultaneous composition, we may think that if the last composition of a specific composition order is non-locking, then every composition is non-locking for all composition orders.
Unfortunately it is not true.
Actually even if the last composition of any order of composition is non-locking there is no guarantee that this order is non-locking at each intermediate step.
However the opposite way is true even if uninformative:
If there is an order of composition where every composition is non-locking then the last composition is non-locking, last composition resulting in the same automaton as the simultaneous composition.
So the condition is better checked after the simultaneous composition or the last individual composition.
\begin{exi}
\begin{figure}
\centering
\input{file_sync.tex}
\caption{A made-up protocol for file synchronisation}
\label{fig:pnls}
\end{figure}
The automaton on Figure \ref{fig:pnls} is a sample protocol for file synchronisation.
The two holes are \(net\) a process to manage network communications, and \(hash\) a process to check files integrity.
The state \(i\) is the initial state.
In that state we want to synchronise the files before working, or work on the local version as a fallback.
If the preferred version is the remote one, we go in state \(d1\) asking the netwotk manager to fetch the lastest version, then ask the integrity checker to validate the files.
If the preferred version is the local one, we go in state \(d2\) asking for the integrity checker to validate the local files, then ask the network manager to push them.
In fallback we use the local version.

@@

\section{Refinement relations in restricted cases}\label{sec:prelref}
The main goal of this section is to explain the different aspects of refinement in restricted cases so that they are not all introduced at once in the real refinement relation.
In particular if we have a family of open automata \(A_i \defobject \OAg[i]\), \(0 \leq i \leq n\), let \(A \defobject \OAg\) be such that \(A = A_0\mdbrk{j_i \mapsto A_i \middle| 1 \leq i \leq n}\).
Recall that there is no relation between the holes of \(A\) and \(A_0\) because \(J = \mpar{\biguplus_{0 \leq i \leq n} J_i} \setminus \mset{j_i}{1 \leq i \leq n}\) is neither subset nor superset of \(J_0\).
So looking at restricted cases where holes are related in a specific manner can help understanding the general case.
Hopefully the intermediate refinement relation introduced can be used as simpler (and less expensive to compute) versions of the real refinement relation in the case the manipulated automata respect some constraints.

We restrict the holes of the related automata so that we can explain the different refinements without their interactions.
In the following definition, \(\triangle \in \mbrc{\subseteq, =, \supseteq}\).
\begin{defi}[Hole-\(\triangle\) simulation]
For two open automata \(A_1 \defobject \OAg[1]\) and \(A_2 \defobject \OAg[2]\), the relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is a hole-\(\triangle\) simulation of \(A_1\) by \(A_2\) if:
\item[1)] \(J_1 \triangle J_2\)
\item[2)] \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\)
\item[3)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)\vspace{-8pt}
\noindent\begin{multline*}
	\mpar{\everymath{\displaystyle}\begin{array}{l}
		\bigsymb{\forall} t_1 \defobject \OTx{1}{}{1}{1} \in \fOT{s_1}, \bigsymb{\exists} \mpar{t_{2x} \defobject \OTx{2}{x}{2x}{2x} \in \fOT{s_2}}^{x \in X}, \\
		\quad \mpar{\forall x \in X, J'_{2x} \cap J_1 = J'_1 \cap J_2} \\[-10pt]
		\nwedge V_1 \uplus V_2 \uplus \fvars{t_1} \vdash R\mpar{s_1, s_2} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\begin{array}{l}
			\alpha_1 = \alpha_{2x} \wedge \bigwedge_\subbox{j \in J'_{2x} \cap J_1} \beta_{1j} = \beta_{2xj} \\[12pt]
			\nwedge g_{2x} \wedge R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}
		\end{array}}
	\end{array}} \\
	\wedge \mpar{V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}}
\end{multline*}
\end{defi}
The third item of the definition has two parts:
The first part is the main requirement, a (not proved equivalent) reordering of the pre-simulation requirement so that it better fits SMT.
Instead of matching an effective transition to another effective transition as in the pre-simulation definition, we match an open transition to a family of covering open-transitions.
The second part is directly the deadlock prevention requirement.

@@

Examples of application of the different refinement relations are given in Appendix \ref{apx:refrel}.
\begin{thm} A hole-\(\triangle\) simulation is a pre-simulation. \end{thm}
Proof is given in Appendix \ref{apx:presim}.
\begin{thm} The hole-\(\triangle\) simulation on automata is a preorder. \end{thm}
The idea for reflexivity is that the related automata are always in the same configuration.
\[ R \defobject \mpar{s, s'} \mapsto \choice{\bigwedge_{v \in V} v_1 = v_2 & \text{if } s = s'\\ \bot & \text{otherwise}} \]
Where \(v_x\) is the renaming to distinguish conflicting automaton variables.
\(R\) is a witness of the reflexivity, the proof is mainly the same as several of the following ones so it will be omitted.

>>>>

\usepackage{xunicode}
\usepackage{fontspec}
\usepackage[hmargin=0.5in,marginparwidth=1.5in,includemp]{geometry}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{unicode-math}
\usepackage{tikz}
\usetikzlibrary{external}
\usepackage[colorlinks]{hyperref}
\AtBeginDocument{\renewcommand\setminus{\smallsetminus}}
\usepackage{polyglossia}
\setmainlanguage{english}

@@

\section{Introduction}
The open automata model was originally designed to give the semantic of open pNets \cite{2007.10770}.
Open pNets are not petri nets but an intermediate representation used in the VerCors project to perform verification \cite{henrio:01252323}.
 Verification is the process of checking that an implementation satisfies a specification.
\Ludo{verification is more general than that (or you need to refine specification): checking typing or absence of deadlocks belong to verification}
Open pNets with the open automaton semantics and equivalence relation introduced in previous work \cite{2007.10770} were sucessfully used to model and verify some properties of applications based on BIP  and GCM \cite{qin:01823507, ameurboulifa:01526055} \Ludo{cite and expand acronym (x2)}.
However equivalence relations are not always sufficient to perform verification, refinement relations are also used. \Ludo{verification/refinement/equivalence needs to be refined and explained}
For instance a usecase where open pNet where equivalence is not sufficient can be found  in an article about algorithms fo weak bisimulation on open automata \cite{wang:03126313}.

\Ludo{choose: either 1) you do not refer to open pnets below and explain informally open automata here (why are they useful, what makes them different from automata), or 2) you want to refer to pnets in the following and you explain informally pnets, uatomata and the relation}

The contributions of this article \Ludo{report?} are the definition of composition for open automata without having to define open pNets, and the introduction of several refinement relations for open automata\marginpar{TODO: expand}.

@@

\Ludo{RAPPEL: pour ecrire une intro: 1) on ecrit les contribs, 2) on explique avant les termeis techniques pour que le lecteur comprenne les contribs, 3) quand on a quasi fini le reste on ecrit le plan. ON peut commencer 1 et 2 un peu tot meme si il faut rediger le tout assez tard: on peut pas ecrire proprement un article sans avoir en tete les contribs de celui ci}

We begin with Section \ref{sec:notations} by giving some notations used throughout this paper.
Then there are several sections dedicated to define all the interesting objects and properties:
In Section \ref{sec:def} we give a clean and up to date definition of open automata.
In Section \ref{sec:comp} we define their composition without requiring the use of open pNets \Ludo{does this mean that previous definitions were using open pnets? this might be clarified in contributions but when you redo a definition you must highlight why a new definition was necessary}.
In Section \ref{sec:proofelts} we adapt several standard properties that we use in the definition of refinement relation from LTS to open automata.
After that Section \ref{sec:prelref} is dedicated to progressively build a refinement relation and explain the elements of this construction.
Section \ref{sec:refinement} introduces and analyses the refinement relation on open automata.
% future work
% conclusion

@@

Families of values, or equivalently maps will be noted \(\mset{i \mapsto x_i}{i \in I}\), \(\mset{i \gets x_i}{i \in I}\) or \(x_i^{i \in I}\).
The latter can only be used if there is a generating expression; for instance \(\mpar{ax}^{x \in \setR}\) represents a scaling function, \(c^{i \in I}\) is a constant function over \(I\).
However \(\mbrc{\alpha \mapsto 1, \beta \mapsto 2, \gamma \mapsto 3}\) has no generating expression and is represented here with the finite version of first notation.
The disjoint union of two maps \(\varphi: I \to X\) and \(\psi: J \to Y\) with \(I \cap J = \emptyset\) is \(\varphi \uplus \psi: I \uplus J \to X \cup Y\).

@@

The terms \(\terms\) is a term algebra\marginpar{reference to come}.
As any term algebra it has constant symbols with arity, variables, and a typing mechanism to distinguish well-formed and ill-formed terms.
The action algebra \(\actions\) is another term algebra.
It can be a subset of the terms.
The formulas \(\formulas\) are at least the first order formulas over \(\terms\) and \(\actions\).
\end{defi}
The term algebras are arbitrary.
The formulas contain at least first order logic terms with an equality relation.
This equality relation is not necessarily a syntactic equality (\(2 + 2 \neq 4\) with a syntactic equality).

@@

Note that in the previous paragraph there are two relations:
The first is between states of two specific automata (\(R \subseteq S_1 \times S_2\)), the second is between automata (\(A_1 \leq A_2\)).
Generally the second is implicit because its definition is always \(\exists R \subseteq S_1 \times S_2, P\mpar{R}\) with \(P\) a property, and only talking about \(R\) and \(P\) is sufficient.
However later on in this article we will define more complex relations on automata so we clarify the distinction now.
A relation of second kind will be called ``relation on automata" whereas a relation of the first kind is called ``relation on configurations" as defined below.
In the work on FH-bisimulation \cite{henrio:01055091} such a relation \(R\) was defined as triples in \(S_1 \times S_2 \times \formulas\) where for \(\mpar{s_1, s_2} \in S_1 \times S_2\) there is a unique \(f \in \formulas\) such that \(\mpar{s_1, s_2, f} \in R\).
This is classically the definition of a function from \(S_1 \times S_2\) to \(\formulas\) so we will define it as such.
\begin{defi}[Relation on configurations]
A relation on configurations of two open automata \(\OAg[1]\) and \(\OAg[2]\) is a function \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\).

@@

The important properties we consider for the relation on states are being a simulation and the prevention of deadlocks as defined in the following definitions.
\begin{defi}[Pre-simulation]\label{def:presim}
For two open automata \(A_1 \defobject \OAg[1]\) and \(A_2 \defobject \OAg[2]\), a relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is a pre-simulation if it satisfies both
\defitem Initial configurations are related: \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\);
\defitem From related states, all out-transitions from \(A_1\) can be simulated in \(A_2\) and their target states are related:
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} t_1 \defobject \OTx{1}{}{1}{1} \in \fOT{s_1}, \forall \sigma: V_1 \uplus V_2 \uplus \fvars{t_1} \to \values, \\
	\mpar{\sigma \vdash R\mpar{s_1, s_2} \wedge g_1} \implies \bigsymb{\exists} t_2 \defobject \OTx{2}{}{2}{2} \in \fOT{s_2}, \exists \nu: \fvars{t_2} \to \values, \\
	\sigma \uplus \nu \vdash \alpha_1 = \alpha_2 \wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} \wedge g_2 \wedge R\mpar{s'_1, s'_2}\psubst{\psi_1 \uplus \psi_2}
\end{multline*}
\end{defi}
This definition translated to open automata the classical definition of simulation \marginpar{cite: Fernandez-Mounier-91-b}.
The last formula means that for every pair of related configurations of the automaton and every possible transitions (\(\vdash g_1\)) from the first automaton, there is a possible transition (\(\vdash g_2\)) such that the produced action matches (\(\alpha_1 = \alpha_2\)), the same holes' action (\(J'_1 \cap J'_2\)) matches (\(\beta_{1j} = \beta_{2j}\)) and the target states are related after variable update. % OR variableS
Apart from holes, this is a natural extension of the notion of simulation on LTS, which is the same definition without variables nor holes nor guards.
In FH-bisimulation, holes with the same name will receive the same open automaton when composed, and the resulting open automaton will still be equivalent.
Here we assume that one of the related open automata can be obtained by composing an open automaton in a hole, open automaton which can have holes itself.
This implies that there is no relation between their holes, for instance composing \(A_i\) in \(j_i\) (\(1 \leq i \leq n\)) of \(A_0\) gives the set of holes \nmm{J \defobject \mpar{\biguplus_\subbox{0 \leq i \leq n} J_i} \setminus \mset{j_i}{1 \leq i \leq n}} that is neither subset nor superset of \(J_0\).
Yet we still assume that in holes with the same name, the same open automaton will be composed.
Hence only the holes in common have to match their actions.
However it is not sufficient to imply properties on the relation on automatons like being a preorder.
It is only a necessary condition for any relation that will be defined with ``simulation" in its name.

@@

% Another way of understanding this predicate is that it is a fixpoint of the union of valuations (function \(f\) defined below) that contains the initial valuation:
% \begin{align*}
	% f\mpar{p} = \mbrc{\mbox{\nmm{s' \mapsto p\mpar{s'} \vee \bigsymb{\bigvee_\subbox{\OTg \,\in\, \fIT{s'}}} p\mpar{s}\psubst{\psi} \wedge g}}} && % TODO: fix that in some way
	% \sigma_0 \vdash \reach{A}\mpar{s_0} &&
	% f\mpar{\reach{A}} = \reach{A}
% \end{align*}
\begin{defi}[Non-locking composition, intuitive definition]
Let \(A_c \defobject \OAg[c]\), \(A_p \defobject \OAg[p]\) and \(j \in J_p\).
The composition \(A \defobject A_p\subst{A_c}{j}\) is non-locking if \(A\) has a reachability predicate satisfying:
From reachable configurations, if there was a possible transition in \(A_p\) then there is a possible transition in \(A_p\subst{A_c}{j}\).

Formally for \(A = \OAg\), it gives
\begin{multline*}
	\forall s \defobject \mpar{s_p, s_c} \in S, \forall \sigma: V \to \values, \mpar{\sigma \vdash \reach{A_p\subst{A_c}{j}}\mpar{s}} \implies \\
	\mpar{\exists t_p \in \fOT{s_p}, \exists \nu_p: \fvars{t_p} \to \values, \sigma \uplus \nu_p \vdash \fguard{t_p}} \implies \\
	\exists t \in \fOT{s}, \exists \nu: \fvars{t} \to \values, \sigma \uplus \nu \vdash \fguard{t}
\end{multline*}
\end{defi}
We can expand the aliases \(S, V\) into \(S_p \times S_c, V_p \uplus V_c\) and transform the last transition \(t\) into a pair \(\mpar{t_p, t}\) where the first transition is the one from \(A_p\) which generated \(t\).
Doing that makes the definition become mildly similar with the deadlock prevention one. % OR slightly
It is not a coincidence because their goal is to caracterise the same kind of compatibility between automata.
Actually it is possible to simplify both definition and make them even more similar.
\begin{defi}[Deadlock prevention, working definition]
A relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is deadlock reducing if it is a pre-simulation and it satisfies the following:
\[ \forall \mpar{s_1, s_2} \in S_1 \times S_2, V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \]
\end{defi}
The requirement here is that from any related configurations, if there is a possible transition in the second automaton (whatever the value of the free variables) then there is a possible transition in the first automaton (the existential quantifier is implicit in the notation \(\vdash\)).
The transition are not matched because the relation on configurations is a pre-simulation and it is sufficient to ensure it as will be shown in the proof.
Apart from the matching transitions this is exactly the same explaination as the intuitive definition.
The expected advantage of this formulation is that a SMT should behave better without the quantifiers.
Indeed the big disjunctions are not quantifiers because in practise the out-transitions are finitary, and so is the expansion of the disjunction. % Check if that construction is correct
\begin{lem}
The intuitive and working definition of deadlock prevention are equivalent.
\end{lem}
\begin{defi}[Non-locking composition, working definition]
The composition \(A \defobject A_p\subst{A_c}{j}\) is a non-locking composition if:
\[ \forall s \defobject \mpar{s_p, s_c} \in S, V \uplus \biguplus_\subbox{t_p \in \fOT{s_p}} \fvars{t_p} \vdash \reach{A}\mpar{s} \wedge \bigvee_\subbox{t_p \in \fOT{s_p}} \fguard{t_p} \implies \bigvee_\subbox{t \in \fOT{s}} \fguard{t} \]
\end{defi}
\begin{lem}
The intuitive and working definition of non-locking composition are equivalent.
\end{lem}
An illustration of both the introduction of a deadlock by composition and the definition of non-locking composition is given in Appendix \ref{apx:lockcomp}.
Proof of both lemma are given in Appendix \ref{apx:lemeqd}.

This new definition of composition can replace the standard composition in the case where only one hole is filled, and also be extended to a simultaneous composition.
For the simultaneous composition, we may think that if the last composition of a specific composition order is non-locking, then every composition is non-locking for all composition order. % orderS
Unfortunately it is not true.
Actually even if the last composition of any order of composition is non-locking there is no guarantee that this order is non-locking at each intermediate step.
However the opposite way is true even if uninformative:
If there is an order of composition where every composition is non-locking then the last composition is non-locking, last composition resulting in the same automaton as the simultaneous composition.
So the condition is better checked after the simultaneous composition or the last individual composition.
\begin{exi}
\begin{figure}
\centering
\input{file_sync.tex}
\caption{A made-up protocol for file synchronisation}
\label{fig:pnls}
\end{figure}
The automaton on Figure \ref{fig:pnls} is a sample protocol for file synchronisation.
The two holes are \(net\) a process to manage network communications, and \(hash\) a process to check files integrity.
The state \(i\) is the initial state.
In that state we want to synchronise the files before working, or work on the local version as a fallback.
If the preferred version is the remote one, we go in state \(d1\) asking the netwotk manager to fetch the lastest version, then ask the integrity checker to validate the files.
If the preferred version is the local one, we go in state \(d2\) asking for the integrity checker to validate the local files, then ask the network manager to push them.
In fallback we use the local version.

@@

\section{Refinement relations in restricted cases}\label{sec:prelref}
The main goal of this section is to explain the different aspects of refinement in restricted cases so that they are not all introduced at once in the real refinement relation.
In particular if we have a family of open automata \(A_i \defobject \OAg[i]\), \(0 \leq i \leq n\), let \(A \defobject \OAg\) be such that \(A = A_0\mdbrk{j_i \mapsto A_i \middle| 1 \leq i \leq n}\).
Recall that there is no relation between the holes of \(A\) and \(A_0\) because \nmm{J = \mpar{\biguplus_{0 \leq i \leq n} J_i} \setminus \mset{j_i}{1 \leq i \leq n}} is neither subset nor superset of \(J_0\).
So looking at restricted cases where holes are related in a specific manner can help understanding the general case.
Hopefully the intermediate refinement relation introduced can be used as simpler (and less expensive to compute) versions of the real refinement relation in the case the manipulated automata respect some constraints.

We restrict the holes of the related automata so that we can explain the different refinements without their interactions.
In the following definition, \(\triangle \in \mbrc{\subseteq, =, \supseteq}\).
\begin{defi}[Hole-\(\triangle\) simulation]
For two open automata \(A_1 \defobject \OAg[1]\) and \(A_2 \defobject \OAg[2]\), the relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is a hole-\(\triangle\) simulation of \(A_1\) by \(A_2\) if:
\defitem \(J_1 \triangle J_2\)
\defitem \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\)
\defitem \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)\vspace{-8pt}
\noindent\begin{multline*}
	\mpar{\everymath{\displaystyle}\begin{array}{l}
		\bigsymb{\forall} t_1 \defobject \OTx{1}{}{1}{1} \in \fOT{s_1}, \bigsymb{\exists} \mpar{t_{2x} \defobject \OTx{2}{x}{2x}{2x} \in \fOT{s_2}}^{x \in X}, \\
		\quad \mpar{\forall x \in X, J'_{2x} \cap J_1 = J'_1 \cap J_2} \\[-10pt]
		\nwedge V_1 \uplus V_2 \uplus \fvars{t_1} \vdash R\mpar{s_1, s_2} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\begin{array}{l}
			\alpha_1 = \alpha_{2x} \wedge \bigwedge_\subbox{j \in J'_{2x} \cap J_1} \beta_{1j} = \beta_{2xj} \\[12pt]
			\nwedge g_{2x} \wedge R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}
		\end{array}}
	\end{array}} \\
	\wedge \mpar{V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}}
\end{multline*}
\end{defi}
The third item of the definition has two parts:
The first part is the main requirement, a (not proved equivalent) reordering of the pre-simulation requirement so that it better fits SMT.
Instead of matching an effective transition to another effective transition as in the pre-simulation definition, we match an open transition to a family of covering open-transitions.
The second part is directly the deadlock prevention requirement.

@@

Examples of application of the different refinement relations are given in Appendix \ref{apx:refrel}.
\begin{thm} A hole-\(\triangle\) simulation is a pre-simulation. \end{thm}
Proof is given in Appendix \ref{apx:presim}.
\begin{thm} The hole-\(\triangle\) simulation on automata is a preorder. \end{thm}
% TODO: give intuition, then define R on one line then renamin on the line after
Reflexivity is immediate with \(R \defobject \mpar{s, s'} \mapsto \choice{\bigwedge_{v \in V} v_1 = v_2 & \text{if } s = s'\\ \bot & \text{otherwise}}\) where \(v_x\) is the renaming to distinguish conflicting automaton variables.
